---
title: "BSUP-684"
author: "Sumeet Tiwari"
date: "27/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objective
Prepare two plots using R (or Python), including the script to produce them in this task.

## Library
```{r library}
library(dplyr)
library(ggplot2)
```

## Input
```{r input}
## Reading the WT and read counts per cell. Files consist ofthree columns: filename, counts, type of reads. 
wt=read.table("~/Desktop/Project/BSUP-684/counts-wt.txt",header=FALSE)
colnames(wt) <- c("filename","reads_count","type")
ko=read.table("~/Desktop/Project/BSUP-684/counts-ko.txt", header=FALSE)
colnames(ko) <- c("filename","reads_count","type")
```

## Histogram for WT and KO sample reads count
Plotting the histogram of reads count from WT and KO samples in individual plots.

```{r}
# Wild-type
ggplot(data=wt, aes(reads_count)) + 
  geom_histogram(bins = 50) +
  labs(title = "Reads count in KO", x="Number of reads", y="count")

# Knocked-out
ggplot(data=ko, aes(reads_count)) + 
  geom_histogram(bins = 50) +
  labs(title = "Reads count in WT", x="Number of reads", y="count")
```

## Histogram WT and KO sample reads count
Plotting the a single histogram for read counts in WT and KO.
```{r}
# Merging the data of WT and KO into one. Label the rows "WT" for wildtpe and "KO" for kocked out in the dataframe
wt_mod=wt[-c(1,3)]
ko_mod=ko[-c(1,3)]
wt_mod$sample=replicate(nrow(wt_mod),"WT")
ko_mod$sample=replicate(nrow(ko_mod), "KO")
wt_ko_df=rbind(wt_mod,ko_mod)


# Plot the historgram by coloring based on the labels "WT" and "KO".
ggplot(data = wt_ko_df, aes(reads_count, color=sample)) +
  geom_histogram(bins = 50, alpha=0.5, position = "dodge", fill="white") +
  labs(title = "Reads count per samples", x="Reads count", y="count")
```

## Filtering samples
The plots are not so clear and it appeared that the data has huge amount of samples with low read count.
```{r}
# Minimum number of reads in a file from wild-type
min(wt_mod$reads_count)
# Maximum number of reads in a file from wild-type
max(wt_mod$reads_count)
# Minimum number of reads in a file from knock-out
min(ko_mod$reads_count)
```


```{r}
# Maximum number of reads in a file from Kock-out
max(ko_mod$reads_count)

# Filtering the samples from both WT and KO based on aribtrary defined threshold on 'reads_sample'.
t=10000 # reads per sample threshold
filter_wt_df=filter(wt_mod, reads_count >= t)
filter_ko_df=filter(ko_mod, reads_count >= t)
filter_wt_ko_df=rbind(filter_wt_df,filter_ko_df)

# Wild-type
ggplot(data=filter_wt_df, aes(reads_count)) + 
  geom_histogram(bins = 50) +
  labs(title = "Reads count in KO", x="Number of reads", y="count")

# Knocked-out
ggplot(data=filter_ko_df, aes(reads_count)) + 
  geom_histogram(bins = 50) +
  labs(title = "Reads count in WT", x="Number of reads", y="count")

# Both Wild-type and Knocked-out 
ggplot(data = filter_wt_ko_df, aes(reads_count, color=sample)) +
  geom_histogram(bins = 50, alpha=0.5, position = "dodge", fill="white") +
  labs(title = "Reads count per samples", x="Reads count", y="count")
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
